---
title: "Assignment for Week 1"
author: "Ahmed Shehata"
date: "March 6, 2016"
output: md_document
---

# 1. Lets Read the file



```{r}
ActivDF <- read.csv('activity.csv')
```


## Load the dplyr library
## and prepare the summary data

```{r}
library(dplyr)

data_summary <- ActivDF[!is.na(ActivDF$steps),c("date","steps")] %>% 
group_by(date) %>%
summarise_each(funs(sum))
```

## 2.Histogram of the total number of steps taken each day  

```{r}
hist(data_summary$steps,col="red",main="Daily Total steps Histogram",xlab="Daily Steps")
```

## 3.Mean and median number of steps taken each day

```{r}
data_summary2 <- ActivDF[!is.na(ActivDF$steps),c("date","steps")] %>% 
group_by(date) %>%
summarise_each(funs(mean))

print(data_summary2)

data_summary3 <- ActivDF[!is.na(ActivDF$steps),c("date","steps")] %>% 
group_by(date) %>%
summarise_each(funs(median))

print(data_summary3)
```

## 4.Time series plot of the average number of steps taken

``` {r}
data_summary4 <- ActivDF[!is.na(ActivDF$steps),c("interval","steps")] %>% 
group_by(interval) %>%
summarise_each(funs(mean))

plot(data_summary4$interval, data_summary4$steps,xlab="Time Series",ylab="Avrage Number of Steps", type = "l")
```

## 5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}
data_summary4[data_summary4$steps == max(data_summary4$steps),]
```

## 6.Code to describe and show a strategy for imputing missing data

```{r}

## assign average of the number of steps for the missing interval
ADF <- ActivDF 

for (interval in data_summary4$interval) 
{
    ADF[is.na(ADF$steps) & ADF$interval == interval   ,1] <- data_summary4[data_summary4$interval==interval,2]
}
```

## 7.Histogram of the total number of steps taken each day after missing values are imputed

```{r}
data_summary5 <- ADF[!is.na(ADF$steps),c("date","steps")] %>% 
group_by(date) %>%
summarise_each(funs(sum))

hist(data_summary5$steps,col="red",main="Daily Total steps Histogram after imputed",xlab="Daily Steps")

```

## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
ADF[,4] <- format(as.Date(ADF$date),"%w")
ADF[ADF$V4 %in% c("1","2","3","4","5"),4] <- "Weekday"
ADF[ADF$V4 %in% c("0","6"),4] <- "Weekend"
names(ADF)[4] <- "DayOfWeek"

library (lattice)

xyplot(steps~interval|factor(DayOfWeek),
       data = ADF,
       type='l',layout=c(1,2),
       xlab='Interval',ylab='Number of Steps')
```

## 9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
